Facility and Time Step Discretization Effects in Fuel Cycle Simulation

### 2015-09-14 paper.tex
restructure sections

\subsection{Motivation and Previous Studies}

@@

\section{Fleet Modeling Overview}

@@

A fleet reactor facility was created for Cyclus.  The flexibility of Cyclus
makes it relatively straight forward to create fleet models for Cyclus that
live anywhere along the spectrum from individual, discrete facilities to
continuous flow fleet facilities.  In order to contrast well with individual,
single-facility granularity modeling, the fleet model created here falls more
toward the aggregate, continuous flow side of the spectrum.

@@

\section{Scenario Description}

The scenario used is a transition from a the current once-through light water
reactor (LWR) fleet to a full sodium fast reactor (SFR) fuel cycle known as
EG-23 in the Department of Energy's (DOE) Advanced Nuclear Fuel Cycle Options
report \cite{wigeland_nuclear_2014}. Scenario and transition details were
patterned after ongoing work by the DOE in their Fuel Cycle Options campaign.

@@

Cyclus version ?.? (TODO:~update~version) was used for this analysis.
Facility models used were the standard ones provided with Cyclus in the
Cycamore library (also version ?.?).  A custom storage facility was used
because Cycamore does not yet provide one.  A custom fleet-based reactor was
also used for both reactor types rather than modeling facilities individually
with the Cycamore reactor primarily for performance reasons. The fleet reactor
has no sophisticated physics functionality and instead just transmutes all
fresh fuel into a fixed spent fuel composition.  A templated Cyclus input file
for the scenario is in Appendix TODO: exp1 input file in appendix. Additional
parameters for the reactor types are shown in Table \ref{tab:reactors}.

@@

\begin{table}[h]
    \centering
    \begin{tabular}{ |r | c c | }
        \hline                       
            & LWR & FR \\
        \hline                       
        Lifetime (yr)         & 80 & 80 \\
        Cycle length (months) & 18 & 14 \\
        Batch size (kg)       & 29565 & 7490 \\
        Batches per core      & 3 & 5 \\
        \hline                       
    \end{tabular}
    \captionsetup{justification=centering}
    \caption[Reactor facility parameters]{Reactor facility parameters.}
    \label{tab:reactors}
\end{table}

The described scenario together with a few perterbations result in four
different cases for investigation:

    \item Monthly time steps with individual reactor modeling
    \item Monthly time steps with fleet reactor modeling
    \item Quarterly (3-month) time steps with individual reactor modeling
    \item Quarterly (3-month) time steps with fleet reactor modeling

A few other small changes to the scenario are also necessary.  For the cases
using quarterly time steps with individual reactors, cycle and outage times
must be adjusted to be a multiple of 3 months.  In order to make the
comparison fair, the cases with monthly time steps are also adjusted to have
identical cycle and outage times.  The invariants preserved with respect to
reactor behavior are shown in Table \ref{tab:invar}.  Table
\ref{tab:reactors} shows the computed/selected configuration for both
reactor types.  In each of the four cases, the initial 100 LWRs are modified
to have a zero length refueling outage with an explicit 900 MWe power
capacity.  This was done to avoid having all the initial LWRs in the
individual reactor cases refuel at the same time.

@@

The parameters for each case were also selected to keep capacity factors for
individual reactors very roughly around 90\%.  The template input file for
each of these cases is in Appendix TODO: exp2 input files in appendix.

@@

>>>>

\subsection{Motivation and Background}

@@

\section{Methodology}

A single fuel cycle transition scenario is used with a few modeling variations
resulting in four different cases:

    \item \emph{Case 1}: Monthly time steps with individual reactor modeling
    \item \emph{Case 2}: Monthly time steps with fleet reactor modeling
    \item \emph{Case 3}: Quarterly (3-month) time steps with individual reactor modeling
    \item \emph{Case 4}: Quarterly (3-month) time steps with fleet reactor modeling

Conducting this analysis involved developing some additional capability
in/around Cyclus.  Three primary pieces to the methodology are:

    \item A fleet-based reactor model.

    \item A simulation scenario with both initial conditions and transition
        details.

    \item Theory and metrics for comparing differences between the four cases.

Because the standard package included with Cyclus does not include a
fleet-based reactor model, one was necessarily developed. A simulation
scenario was created that is roughly based on prior and ongoing work in the
Department of Energy (DOE).  A few Phenomena related to the different modeling
choices in each of the four cases are identified and used as a basis for
comparison.  Each of these three things is described in more detail in the
following sections.

@@

A fleet reactor facility was created for Cyclus.  The flexibility of Cyclus
makes it relatively straight forward to create plug-in models for Cyclus that
live anywhere along the spectrum from individual, discrete facilities to
continuous flow fleet facilities.  In order to contrast well with individual,
single-facility granularity modeling, the fleet model created here falls more
toward the aggregate, continuous flow side of the spectrum.

@@

\subsection{Scenario Description}

The scenario used is a transition from a the current once-through light water
reactor (LWR) fleet to a full sodium fast reactor (SFR) fuel cycle known as
EG-23 in the DOE's Advanced Nuclear Fuel Cycle Options report
\cite{wigeland_nuclear_2014}. Scenario and transition details were patterned
after ongoing work by the DOE in their Fuel Cycle Options campaign.

@@

Cyclus version ?.? (TODO:~update~version) was used for this analysis.
Facility models used were the standard ones provided with Cyclus in the
Cycamore library (also version ?.?).  A custom storage facility was used
because Cycamore does not yet provide one.  The standard cycamore reactor was
used for both reactor types (LWR and SFR) in cases 1 and 3, and the custom
fleet-based reactor described earlier also used in cases 2 and 4.
Additional parameters for the reactor types are shown in Table
\ref{tab:reactors}.

@@

\begin{table}[h]
    \centering
    \begin{tabular}{ |r | c c | }
        \hline                       
            & LWR & FR \\
        \hline                       
        Lifetime (yr)         & 80 & 80 \\
        Cycle length (months) & 18 & 14 \\
        Batch size (kg)       & 29565 & 7490 \\
        Batches per core      & 3 & 5 \\
        \hline                       
    \end{tabular}
    \captionsetup{justification=centering}
    \caption[Reactor facility parameters]{Reactor facility parameters. The cycle length includes the refueling outage.}
    \label{tab:reactors}
\end{table}

The described scenario is used as closely as possible in each of the four
cases.  However, a few other small case-dependent changes to the scenario
relating to reactor configuration are necessary.  For the cases using
quarterly time steps with individual reactors, cycle and outage times must be
adjusted to be a multiple of 3 months.  In order to make the comparison more
fair, the cases with monthly time steps are also adjusted to have identical
cycle and outage times.  The invariants preserved with respect to reactor
behavior are shown in Table \ref{tab:invar}.  Table \ref{tab:reactors} shows
the computed/selected configuration for both reactor types for all four cases.

@@

In each of the four cases, the initial 100 LWRs are modified to have a zero
length refueling outage with an explicit 900 MWe power capacity.  This was
done to avoid having all the initial LWRs in the individual reactor cases
refuel at the same time.  The parameters for each case were also selected to
keep capacity factors for individual reactors very roughly around 90\%.

@@

The input files and other assets used in the analysis for each of the four
cases can be found in TODO: cite figshare tarball.

\subsection{Modeling Effects}
